---
title: "in-class 10/20"
author: "Michelle"
date: "10/20/2020"
output: 
  html_document: 
   toc: TRUE
   toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Here is a demonstration of how to use R Markdown!


## Making a new code chunk

The shortcut key (on windows) is Ctrl-Alt-i

If I press this key combo, Rstudio will insert a new code chunk!

If I want to, I can give the chunk a name and it will appear in the navigation pane.

```{r how to make a new chunk}

```

## Reading a csv file

I'm going to read some data from a csv (comma-separated-values) file

I can use read.csv() along with the path to the file to read the data into a data.frame.

Here's an example:


```{r read_birds_data}
require(here)

dat_birds = read.csv(here("data", "hab.sta.csv"))
dat_birds <- read.csv(here("data", "hab.sta.csv"))


```





## Previewing the contents of a data.frame

I can use the head() function to print the first few lines of a data.frame.
Here's an example with the birds data:

```{r head function}
head(dat_birds)

```



##Demo Plot

I can use hist() to create a histogram:

When I plot aspect, the last column is weird..


```{r simple plot}
hist(dat_birds$aspect)

```


## How to customize a histogram

I can title using the argument main =
I can customize the bins using the argument breaks = 

```{r custom histogram 1}

hist(
  dat_birds$aspect, 
  main = "Histogram of Study Site Aspect",
  breaks = 36)
```

